---
title: "codebook"
author: "Nazmi Anik"
date: "Sunday, July 26, 2015"
output: html_document
---

# Brief Summary

Data used in the code is collected from the accelerometers from the Samsung Galaxy S smartphone. 
Subjects were asked to lay down, sit, stand, walk upstairs, walk downstairs and walk with the smart phone.

# Methodology

The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ.

These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz.

Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz.

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ).

Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag).

Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals).

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.

# Data Manipulation

The data listed in the tidydata.txt file was filtered from 561 different measurements to include only ones with
standard deviations and means of the measurements.

This is how the filtering was done:

```{r eval = FALSE}
    ## look for -mean() or -std()
    selectMeasurements <- grep("-(mean|std)\\(\\)", features[, 2])
    
    ## grab the labels
    selectLabels <- features[selectMeasurements,2]
    
    ## subset the chosen measurements
    selectedData <- mergedxData[,selectMeasurements]
    
    ## add the y and subject data columns to the table
    finalData <- cbind(mergedsubjectData, mergedyData, selectedData)
```

This way we ended up with 68 variables.

Next the data was tidied as follows.

1. The column names were updated:

```{r eval = FALSE}
    ## replace finalData table's activity id column with activity name
    finalData[,2] <- activityLabels[finalData[,2], 2]
    
    
    # Appropriately label the data set with descriptive variable names. 
    
    
    names(finalData)<-gsub("-", ".", names(finalData))
    names(finalData)<-gsub("subjectId", "SubjectId", names(finalData))
    names(finalData)<-gsub("activityId", "Activity", names(finalData))
    names(finalData)<-gsub("^t", "Time", names(finalData))
    names(finalData)<-gsub("^f", "Frequency", names(finalData))
    names(finalData)<-gsub("Acc", "Accelerometer", names(finalData))
    names(finalData)<-gsub("Gyro", "Gyroscope", names(finalData))
    names(finalData)<-gsub("Mag", "Magnitude", names(finalData))
    names(finalData)<-gsub("BodyBody", "Body", names(finalData))
    names(finalData)<-gsub("std", "StandardDeviation", names(finalData))
    names(finalData)<-gsub("mean", "Mean", names(finalData))
    names(finalData)<-gsub("\\(|\\)", "", names(finalData))
```

2. The data was narrowed (melted):

```{r eval = FALSE}
    # melt the data set and make it narrow and tidy
    library(reshape2)
    finalData.melted <- melt(finalData, id = c("SubjectId", "Activity"))
```

Finally, the mean of every data for each subject-activity pair was calculated and written in a tidy data file.

```{r eval = FALSE}
    ## cast/summarize the data with the mean function, wide and tidy
    finalData.mean <- dcast(finalData.melted, SubjectId + Activity ~ variable, mean)
    
    ## write the wide tidy table to a txt file
    write.table(finalData.mean, file = ".\\tidydata.txt", row.names = FALSE)
```

# Variable Descriptions

## General Variables

SubjectID: Identifier for each subject.

Activity: Describes the activity performed by the subject.

## Measurements

TimeBodyAccelerometer.Mean.X

TimeBodyAccelerometer.Mean.Y

TimeBodyAccelerometer.Mean.Z

TimeBodyAccelerometer.StandardDeviation.X

TimeBodyAccelerometer.StandardDeviation.Y

TimeBodyAccelerometer.StandardDeviation.Z

TimeGravityAccelerometer.Mean.X

TimeGravityAccelerometer.Mean.Y

TimeGravityAccelerometer.Mean.Z

TimeGravityAccelerometer.StandardDeviation.X

TimeGravityAccelerometer.StandardDeviation.Y

TimeGravityAccelerometer.StandardDeviation.Z

TimeBodyAccelerometerJerk.Mean.X

TimeBodyAccelerometerJerk.Mean.Y

TimeBodyAccelerometerJerk.Mean.Z

TimeBodyAccelerometerJerk.StandardDeviation.X

TimeBodyAccelerometerJerk.StandardDeviation.Y

TimeBodyAccelerometerJerk.StandardDeviation.Z

TimeBodyGyroscope.Mean.X

TimeBodyGyroscope.Mean.Y

TimeBodyGyroscope.Mean.Z

TimeBodyGyroscope.StandardDeviation.X

TimeBodyGyroscope.StandardDeviation.Y

TimeBodyGyroscope.StandardDeviation.Z

TimeBodyGyroscopeJerk.Mean.X

TimeBodyGyroscopeJerk.Mean.Y

TimeBodyGyroscopeJerk.Mean.Z

TimeBodyGyroscopeJerk.StandardDeviation.X

TimeBodyGyroscopeJerk.StandardDeviation.Y

TimeBodyGyroscopeJerk.StandardDeviation.Z

TimeBodyAccelerometerMagnitude.Mean

TimeBodyAccelerometerMagnitude.StandardDeviation

TimeGravityAccelerometerMagnitude.Mean

TimeGravityAccelerometerMagnitude.StandardDeviation

TimeBodyAccelerometerJerkMagnitude.Mean

TimeBodyAccelerometerJerkMagnitude.StandardDeviation

TimeBodyGyroscopeMagnitude.Mean

TimeBodyGyroscopeMagnitude.StandardDeviation

TimeBodyGyroscopeJerkMagnitude.Mean

TimeBodyGyroscopeJerkMagnitude.StandardDeviation

FrequencyBodyAccelerometer.Mean.X

FrequencyBodyAccelerometer.Mean.Y

FrequencyBodyAccelerometer.Mean.Z

FrequencyBodyAccelerometer.StandardDeviation.X

FrequencyBodyAccelerometer.StandardDeviation.Y

FrequencyBodyAccelerometer.StandardDeviation.Z

FrequencyBodyAccelerometerJerk.Mean.X

FrequencyBodyAccelerometerJerk.Mean.Y

FrequencyBodyAccelerometerJerk.Mean.Z

FrequencyBodyAccelerometerJerk.StandardDeviation.X

FrequencyBodyAccelerometerJerk.StandardDeviation.Y

FrequencyBodyAccelerometerJerk.StandardDeviation.Z

FrequencyBodyGyroscope.Mean.X

FrequencyBodyGyroscope.Mean.Y

FrequencyBodyGyroscope.Mean.Z

FrequencyBodyGyroscope.StandardDeviation.X

FrequencyBodyGyroscope.StandardDeviation.Y

FrequencyBodyGyroscope.StandardDeviation.Z

FrequencyBodyAccelerometerMagnitude.Mean

FrequencyBodyAccelerometerMagnitude.StandardDeviation

FrequencyBodyAccelerometerJerkMagnitude.Mean

FrequencyBodyAccelerometerJerkMagnitude.StandardDeviation

FrequencyBodyGyroscopeMagnitude.Mean

FrequencyBodyGyroscopeMagnitude.StandardDeviation

FrequencyBodyGyroscopeJerkMagnitude.Mean

FrequencyBodyGyroscopeJerkMagnitude.StandardDeviation



